Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-09-20 paper/main.tex
move morse potential to supplement

\begin{equation}
    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
    \label{eq:cell-force-external}
\end{equation}

The specific choice of interaction potential plays of course a crucial role for the dynamics of the
agents.
But we cannot expect to simply know the shape and parameters of this potential.
Thus, we would like to experiment with various interaction potentials in order to quantify their
ability to produce accurate results.
We utilize two distinct potentials that are both able to capture the essence of our interactions by
providing a short-range repulsive part and a long-ranged attractive component.

\paragraph{Morse Potential}
The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
potential.
It was originally designed to describe the interaction of diatomic molecules but has since found use
in other areas~\cite{Breitwieser2021}.
It contains a repulsive and attractive part and is bound from above which makes it particularly
suitable for numerical simulations since unbound potentials can lead to large numerical deviations
thus compromising the prediction of the model.
The potential is of the form
\begin{equation}
    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
\end{equation}
and has 3 parameters.
The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
$r\geq\xi$.

\paragraph{Mie Potential}
The Mie~\cite{Mie1903} potential is the generalization of the well-known
Lennard-Jones~\cite{Jones1924} potential.
It was also designed to describe the interactions of molecules but due to the flexible nature of its
parameters, can be easily adapted to other situations.
It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
and not bound from above.
In our simulation, we enforce an upper bound to guarantee numerical regularity.
It has 4 parameters and is given by
\begin{align}
    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
        \left(\frac{\sigma}{r}\right)^m\right]\\
    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
\end{align}
where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
the potential.
The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.

@@

>>>>

\begin{equation}
    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
\end{equation}

The Mie~\cite{Mie1903} potential is the generalization of the well-known
Lennard-Jones~\cite{Jones1924} potential.
It was also designed to describe the interactions of molecules but due to the flexible nature of its
parameters, can be easily adapted to other situations.
It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
and not bound from above.
In our simulation, we enforce an upper bound to guarantee numerical regularity.
It has 4 parameters and is given by
\begin{align}
    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
        \left(\frac{\sigma}{r}\right)^m\right]\\
    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
\end{align}
where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
the potential.
The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.

The specific choice of interaction potential plays a crucial role for the dynamics of the agents.
Since we cannot expect to simply know the correct shape and parameters of this potential, it is
desirable to experiment with various interaction potentials in order to quantify their
ability to produce accurate results.
While this study, only relies on the Mie Potential, our python package \texttt{cr\_mech\_coli}
also implements the popular Morse potential~\cite{Morse1929} to model physical interactions.

@@

%###################################################################################################
\subsection{Alternative Interaction Potential - Morse Potential}
The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
potential.
It was originally designed to describe the interaction of diatomic molecules but has since found use
in other areas~\cite{Breitwieser2021}.
It contains a repulsive and attractive part and is bound from above which makes it particularly
suitable for numerical simulations since unbound potentials can lead to large numerical deviations
thus compromising the prediction of the model.
The potential is of the form
\begin{equation}
    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
\end{equation}
and has 3 parameters.
The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
$r\geq\xi$.
